---
title: "R Bootcamp Intro Exercises"
author: "Emily Hamant"
date: "Last modified 2026-01-27"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Mathematical operations 

```{r}
2 + 5 
2*5
2/5
2^5
```

# Creating objects 

```{r}
x <- 3 
y <- 3 
x + y 
x * y
x.2 <- "hello"
```

# Types of Objects 

## Scalar

only composed of one item. 

```{r}
x <- 3 
scalar <- 2 + 2 
dairy.bar <- "yes please!"
```

## Vectors

vector has to be the same type of data 

```{r}
v1 <- c(2,10,5)

v2 <- c("Prof", "Courtney", "Davis")

length(v1)
```

## Matrices 

```{r}
x1 <- c(1,3,3)

x2 <- c(4,5,6)

mat1 <- cbind(x1, x2) #binds items by column. They make the vectors turn into columns. First column is 1,3,3

mat1

mat2 <- rbind(x1, x2) #binds items by rows. first row is 1,3,3,

mat2

mat3 <- matrix(c(1,2,3,4,5,6), nrow = 3, ncol = 2, byrow = TRUE) # want matrix with 3 rows, 2 columns. fill by row. 

mat3

dim(mat3)

mat4 <- matrix(c(1,2,3,4,5,6), nrow = 3, ncol = 2, byrow = FALSE) # fills by columns

mat4
```

## Indexing 

Choose particular components of matrices/vectors

```{r}
v3 <- 1:15 # this is a vector
v3[2]
3v3[2,3,7] # this makes r think that there are multiple dimensions, but there is only one. 
v3[c(2.3.7)]
v3[-2] #exclude 2 


mat3[-2, 1]

## remove everything but the element in row 2, column 1 from mat 3 

mat3

mat3[-c(1,3), -2]

```

## Subsetting Dataframes 

```{r}
df2$month
df2[ ,1]
df2[ ,”month”]
df2[2,1]
df2$month[2]

df2$flavor <- c(“Vanilla”, “Chocolate”, “Mint”) # new column
```

## Classes 

```{r}
class(df2$month)
class(df2$DB_appeal)
is.numeric(df2$month)
is.numeric(df2$DB_appeal)
```


# Boot Camp Practice 1 

1. 

test is a character, because v2 is a character. 

```{r}
test <- c(v1, v2)

class(test)
```
2. Remove 2, 4, 5 components of test 

```{r}
test[c(-2,-4,-5)]
```

## Using which

```{r}
x <- 1:15

which(x == 12) # are there any values that equal 12 already? This will give the element that is equal to 12

which( x!= 10) #shows the elements that are not equal to 12

which( x == 4 | x == 5) # | is the sign for "or"

which( x == 4 & x == 5 ) # & is the sign for "and"

worldfloras[which(worldfloras$Population >100), ] # shows the rows from the worldflora dataset with all the associated columns

```


## Using apply

X = subsetting out columns 2-5, this is the data we are applying the function to. 

MARGIN = this is the dimension we are applying this function to. 

FUN = the function we are applying

```{r}

apply(X = worldfloras[,c(2:5)], MARGIN = 2, FUN = mean) 
```

There are a lot of types of apply functions 

ex: tapply 

applies a function across a group ID that you specify 

X = data we are applying this to 

INDEX = group ID we are applying the function to 

ex: wanting to know the mean flora by continent



```{r}
tapply( X = worldfloras$Flora, INDEX = worldfloras$Continent, FUN = mean)
```

length returns the number of elements in a vector
